Secure Composition of Robust and Optimising Compilers

### 2023-08-11 tech/main.tex
Add splitting

\begin{gather*}
  \begin{aligned}
    \mi{(Expressions)}~&\src{\expr}\bnfdef\ \src{x} \mid \src{\valueexpr} \mid \src{\lbinop{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lget{x}{\expr}} \mid \src{\llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]}} \mid \src{\ldelete{x}} \mid \src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}} \\
                       & \mid \src{\lpair{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lproja{\expr}} \mid \src{\lprojb{\expr}} \mid \src{\llet{x}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lreturn{\expr}} \mid \src{\lcall{g}{\expr}} \mid \src{\lifz{\expr}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\labort}
  \end{aligned}
  \\
  \begin{aligned}
  \mi{(Types)}~&\src{\type}\bnfdef\src{\natt} \mid \src{\type[_{1}]} \times \src{\type[_{2}]}
  &
  \mi{(Functions)}~&\src{\function}\bnfdef\ \src{\lfunction{\textit{foo}}{x}{e}}
  &
\mi{(Libraries)}~&\src{\library}\bnfdef\ \src{\hole{\cdot}} \mid \src{\function},\src{\library}
  \end{aligned}\\
  \begin{aligned}
  \mi{(Component~Names)}~&\src{\commlib}\bnfdef\ \src{\hole{\cdot}} \mid \src{\text{foo}},\src{\commlib}
  &
  \mi{(Programs)}~&\src{\prog{\library_{\ctx}}{\library_{\comp}}}
  \end{aligned}
\end{gather*}

@@

    \mi{(Control~Tags)}~&\src{\sandboxtag}\bnfdef\src{\ctx} \mid \src{comp}
    &
    \mi{(Communication~Tags)}~&\src{\comm}\bnfdef\ \src{\ctxtocomp} \mid\ \src{\comptoctx} \mid \src{\nocomm}
    &
    \mi{(Poison~Tags)}~&\src{\poison}\bnfdef \src{\poisoned} \mid \src{\poisonless}
    \mi{(Continuation~Stacks)}~&\src{\ectxstack}\bnfdef\src{\hole{\cdot}} \mid \src{\tup{K}{g}},\src{\ectxstack}
    &
    \mi{(Control~Flow~States)}~&\src{\cfstate}\bnfdef\src{\tup{\library}{\ectxstack}}
  	\\
    \mi{(Locations)}~&\src{\loc}\in\nat
    &
    \mi{(Substitutions)}~&\src{\substvar}\bnfdef \src{\hole{\cdot}} \mid \src{\subst{\valueexpr}{x}},\src{\gamma}
  	\\
    \mi{(Stores)}~&\src{\Delta}\bnfdef\src{\hole{\cdot}} \mid \src{x\mapsto\tup{\loc}{\sandboxtag,\poison,n}},\src{\Delta}
    &
    \mi{(Heaps)}~&\src{H}\bnfdef\src{\hole{\cdot}} | \src{\valueexpr},\src{H}
    \mi{(Memory~States)}~&\src{\memstate}\bnfdef\src{\tup{H^{\ctx}}{H^{\comp};\Delta}}
    &
    \mi{(States)}~&\src{\statevar}\bnfdef\src{\tup{\cfstate}{\sandboxtag,\memstate}}
    &
    \mi{(Runtime~Terms)}~&\src{\runtimetermvar}\bnfdef\src{\rtt{\statevar}{\expr}} \mid \src{\lightning}
Throughout the paper, the substitution notation is also used to update entries in states $\statevar$.

@@

@@

>>>>

\begin{gather*}
  \begin{aligned}
    \mi{(Expressions)}~&\src{\expr}\bnfdef\ \src{x} \mid \src{\valueexpr} \mid \src{\lbinop{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lget{x}{\expr}} \mid \src{\lnew{x}{\expr[_{1}]}{\expr[_{2}]}{\expr[_{3}]}} \mid \src{\ldelete{x}} \mid \src{\lset{x}{\expr[_{1}]}{\expr[_{2}]}} \\
                       & \mid \src{\lpair{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lproja{\expr}} \mid \src{\lprojb{\expr}} \mid \src{\llet{x}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\lreturn{\expr}} \mid \src{\lcall{g}{\expr}} \mid \src{\lifz{\expr}{\expr[_{1}]}{\expr[_{2}]}} \mid \src{\labort}
  \end{aligned}
  \\
  \begin{aligned}
  \mi{(Types)}~&\src{\type}\bnfdef\src{\natt} \mid \src{\ptr} \mid \src{\wptr} \mid \src{\type[_{1}] \times \type[_{2}]}
  &
  \mi{(Type~Enviroments)}~&\src{\typenv}\bnfdef\src{\hole{\cdot}} \mid \src{x:\type,\typenv} \mid \src{x:\type[_1]\to\type[_2],\typenv}
  \end{aligned}\\
  \begin{aligned}
  \mi{(Functions)}~&\src{\function}\bnfdef\ \src{\lfunction{\textit{foo}}{x}{e}}
  &
\mi{(Libraries)}~&\src{\library}\bnfdef\ \src{\hole{\cdot}} \mid \src{\function},\src{\library}
  \end{aligned}\\
  \begin{aligned}
  \mi{(Component~Names)}~&\src{\commlib}\bnfdef\ \src{\hole{\cdot}} \mid \src{\text{foo}},\src{\commlib}
  &
  \mi{(Programs)}~&\src{\prog{\library_{\ctx}}{\library_{\comp}}}
  \end{aligned}
\end{gather*}

@@

\begin{gather*}
  \begin{aligned}
    \mi{(Control~Tags)}~&\src{\sandboxtag}\bnfdef\src{\ctx} \mid \src{comp}
    &
    \mi{(Communication~Tags)}~&\src{\comm}\bnfdef\ \src{\ctxtocomp} \mid\ \src{\comptoctx} \mid \src{\nocomm}
    &
    \mi{(Poison~Tags)}~&\src{\poison}\bnfdef \src{\poisoned} \mid \src{\poisonless}
  \end{aligned}\\
  \begin{aligned}
    \mi{(Continuation~Stacks)}~&\src{\ectxstack}\bnfdef\src{\hole{\cdot}} \mid \src{\tup{K}{g}},\src{\ectxstack}
    &
    \mi{(Control~Flow~States)}~&\src{\cfstate}\bnfdef\src{\tup{\library}{\ectxstack}}
  	\\
    \mi{(Locations)}~&\src{\loc}\in\nat
    &
    \mi{(Substitutions)}~&\src{\substvar}\bnfdef \src{\hole{\cdot}} \mid \src{\subst{\valueexpr}{x}},\src{\gamma}
  	\\
    \mi{(Stores)}~&\src{\Delta}\bnfdef\src{\hole{\cdot}} \mid \src{x\mapsto\tup{\loc}{\sandboxtag,\poison,n}},\src{\Delta}
    &
    \mi{(Heaps)}~&\src{H}\bnfdef\src{\hole{\cdot}} | \src{\valueexpr},\src{H}
  \end{aligned}\\
  \begin{aligned}
    \mi{(Memory~States)}~&\src{\memstate}\bnfdef\src{\tup{H^{\ctx}}{H^{\comp};\Delta}}
    &
    \mi{(States)}~&\src{\statevar}\bnfdef\src{\tup{\cfstate}{\sandboxtag,\memstate}}
    &
    \mi{(Runtime~Terms)}~&\src{\runtimetermvar}\bnfdef\src{\rtt{\statevar}{\expr}} \mid \src{\lightning}
  \end{aligned}
\end{gather*}
Throughout the paper, the substitution notation is also used to update entries in states $\statevar$.

\begin{center}
  \judgbox{\noownedptr{\src{\type}}}{,,Type $\src{\type}$ does not entail an owned pointer.''}
  %
  \typerule{$\src{\Ltms}-\text{noownedptr}-\src{\natt}$}{
  }{
    \noownedptr{\src{\natt}}
  }{ap:ltms-noownedptr-natt}
  %
  \typerule{$\src{\Ltms}-\text{noownedptr}-\src{\wptr}$}{
  }{
    \noownedptr{\src{\wptr}}
  }{ap:ltms-noownedptr-wptr}
  %
  \typerule{$\src{\Ltms}-\text{noownedptr}-\text{pair}$}{
    \noownedptr{\src{\type[_1]}} &
    \noownedptr{\src{\type[_2]}}
  }{
    \noownedptr{\src{\type[_1]\times\type[_2]}}
  }{ap:ltms-noownedptr-pair}
\end{center}
\begin{center}
  \judgbox{\noownedptr{\src{\typenv}}}{,,$\src{\typenv}$ does not contain any owned pointers.''}
  %
  \typerule{$\src{\Ltms}-\text{noownedptr}-\text{empty}$}{
  }{
    \noownedptr{\src{\hole{\cdot}}}
  }{ap:ltms-noownedptr-empty}
  %
  \typerule{$\src{\Ltms}-\text{noownedptr}-\text{cons}$}{
    \noownedptr{\src{\type}} &
    \noownedptr{\src{\typenv}}
  }{
    \noownedptr{\src{x:\type,\typenv}}
  }{ap:ltms-noownedptr-cons}
\end{center}
\begin{center}
  \judgbox{\src{\typenv}\equiv\typenvsplit{\src{\typenv[']}}{\src{\typenv['']}}}{,,Split environment $\src{\typenv}$ into $\src{\typenv[']}$ and $\src{\typenv['']}$.''}
  %
  \typerule{$\src{\Ltms}-$split-empty}{
  }{
    \src{\hole{\cdot}} \equiv \typenvsplit{\src{\hole{\cdot}}}{\src{\hole{\cdot}}}
  }{ap:ltms-split-empty}
  %
  \typerule{$\src{\Ltms}-$split-emptyL}{
  }{
    \src{\typenv} \equiv \typenvsplit{\src{\hole{\cdot}}}{\src{\typenv}}
  }{ap:ltms-split-emptyL}
  %
  \typerule{$\src{\Ltms}-$split-emptyR}{
  }{
    \src{\typenv} \equiv \typenvsplit{\src{\typenv}}{\src{\hole{\cdot}}}
  }{ap:ltms-split-emptyR}
  %
  \typerule{$\src{\Ltms}$-split-$\natt$}{
    \src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}
  }{
    \src{x:\natt},\src{\typenv} \equiv \typenvsplit{\src{x:\natt},\src{\typenv[_1]}}{\src{x:\natt},\src{\typenv[_2]}}
  }{ap:ltms-split-natt}
  %
  \typerule{$\src{\Ltms}$-split-wptr}{
    \src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}
  }{
    \src{x:\wptr},\src{\typenv} \equiv \typenvsplit{\src{x:\wptr},\src{\typenv[_1]}}{\src{x:\wptr},\src{\typenv[_2]}}
  }{ap:ltms-split-wptr}
  %
  \typerule{$\src{\Ltms}$-split-pair}{
    \noownedptr{\src{\type[_1]\times\type[_2]}} &
    \src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}
  }{
    \src{x:\type[_1]\times\type[_2]},\src{\typenv} \equiv \typenvsplit{\src{x:\type[_1]},\src{\typenv[_1]}}{\src{x:\type[_2]},\src{\typenv[_2]}}
  }{ap:ltms-split-pair}
  %
  \typerule{$\src{\Ltms}$-split-ptrL}{
    \src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}
  }{
    \src{x:\ptr},\src{\typenv} \equiv \typenvsplit{\src{x:\ptr},\src{\typenv[_1]}}{\src{\typenv[_2]}}
  }{ap:ltms-split-ptrL}
  %
  \typerule{$\src{\Ltms}$-split-ptrR}{
    \src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}
  }{
    \src{x:\ptr},\src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{x:\ptr},\src{\typenv[_2]}}
  }{ap:ltms-split-ptrR}
  %
  \typerule{$\src{\Ltms}$-split-arrow}{
    \noownedptr{\src{\type[_1]}} &
    \noownedptr{\src{\type[_2]}} &
    \src{\typenv} \equiv \typenvsplit{\src{\typenv[_1]}}{\src{\typenv[_2]}}
  }{
  \src{foo:\type[_1]\to\type[_2]},\src{\typenv} \equiv \typenvsplit{\src{foo:\type[_1]\to\type[_2]},\src{\Gamma_1}}{\src{foo:\type[_1]\to\type[_2]},\src{\Gamma_2}}
  }{ap:ltms-split-arrow}
\end{center}

@@

  }{ap:ltms-e-proj-b}

@@

  \judgbox{\estepton{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\trace}}}{,,$\src{\runtimetermvar}$ performs $n$ steps to $\src{\runtimetermvar[']}$ emitting trace $\src{\trace}$.''}
  \typerule{$\src{\Ltms}-\text{estepn}-\text{refl}$}{
    \estepton[0]{\src{\runtimetermvar}}{\src{\runtimetermvar}}{\src{\hole{\cdot}}}
  }{ap:ltms-estepn-refl}
  \typerule{$\src{\Ltms}-\text{estepn}-\text{trans}-\text{important}$}{
    \src{\event}\not=\src{\emptyevent} &
    \estepto[]{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\event}} &
    \estepton[n]{\src{\runtimetermvar[']}}{\src{\runtimetermvar['']}}{\src{\trace}}
    \estepton[n+1]{\src{\runtimetermvar}}{\src{\runtimetermvar['']}}{\src{\event\cdot\trace}}
  }{ap:ltms-estepn-trans-important}
  \typerule{$\src{\Ltms}-\text{estepn}-\text{trans}-\text{unimportant}$}{
    \estepto[]{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\emptyevent}} &
    \estepton[n]{\src{\runtimetermvar[']}}{\src{\runtimetermvar['']}}{\src{\trace}}
    \estepton[n+1]{\src{\runtimetermvar}}{\src{\runtimetermvar['']}}{\src{\trace}}
  }{ap:ltms-estepn-trans-unimportant}
  \judgbox{\isfinal{\src{\runtimetermvar}}}{,,Runtime term $\src{\runtimetermvar}$ represents a final state.''}
  \typerule{$\src{\Ltms}-\src{\lightning}-\text{is}-\text{final}$}{
    \isfinal{\src{\lightning}}
  }{ap:ltms-stuck-is-final}
  \typerule{$\src{\Ltms}-\src{\valueexpr}-\text{is}-\text{final}$}{
    \isfinal{\src{\rtt{\statevar}{\valueexpr}}}
  }{ap:ltms-value-is-final}
  \judgbox{\estepto{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\trace}}}{,,$\src{\runtimetermvar}$ steps to final $\src{\runtimetermvar[']}$ emitting trace $\src{\trace}$.''}
  \typerule{$\src{\Ltms}-\text{esteps}-\text{refl}$}{
    \isfinal{\src{\runtimetermvar}}
    \estepto{\src{\runtimetermvar}}{\src{\runtimetermvar}}{\src{\hole{\cdot}}}
  }{ap:ltms-esteps-refl}
  \typerule{$\src{\Ltms}-\text{esteps}-\text{trans}-\text{important}$}{
    \src{\event}\not=\src{\emptyevent} &
    \estepto[]{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\event}} &
    \estepto{\src{\runtimetermvar[']}}{\src{\runtimetermvar['']}}{\src{\trace}}
    \estepto{\src{\runtimetermvar}}{\src{\runtimetermvar['']}}{\src{\event\cdot\trace}}
  }{ap:ltms-esteps-trans-important}
  \typerule{$\src{\Ltms}-\text{esteps}-\text{trans}-\text{unimportant}$}{
    \estepto[]{\src{\runtimetermvar}}{\src{\runtimetermvar[']}}{\src{\emptyevent}} &
    \estepto{\src{\runtimetermvar[']}}{\src{\runtimetermvar['']}}{\src{\trace}}
    \estepto{\src{\runtimetermvar}}{\src{\runtimetermvar['']}}{\src{\trace}}
  }{ap:ltms-esteps-trans-unimportant}
